The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-10-02 manuscript.Rmd
a few more edits from PS

Several factors may contribute for species to be included in the regional lists but to not have records in biodiversity databases. The first category of biases is linked to the likelihood of having specimens for these species in museum collections. For instance, rare species may never have been collected or have only been collected a few times, and the corresponding records have not been digitized yet. Species that are difficult to preserve because they are large, very small, or fragile, are deposited in museums less frequently. The other category of biases is associated with data quality for these records. Some of these species may be in the databases but do not have geographic coordinates associated with them, and therefore would not be included in the results of our query. While it is difficult to assess biases associated with the likelihood of preservation of a species, we investigated how missing geographic coordinates are affecting the discrepancies between the checklists and the databases.

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio, we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for the ones that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_usa*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_usa, big.mark=",")`) fall within US waters. Additional digitization of collections, including the capture of their geographic coordinates information, are needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems.

Two factors most likely contribute to explaining why some species are in databases but are not included in the regional the checklists: (1) the species have been documented for the first time since the publication of the checklist, (2) the species records were not accessible or known to the authors.

Recently recorded species contribute modestly to this discrepancy for the Gulf of Mexico, as `r sum(recently_collected$gom_in_dbs$min_year > 2001)` species (`r format_output( sum(recently_collected$gom_in_dbs$min_year > 2001)/n_distinct(total_gom_records$worms_valid_name) * 100)`% of the total species richness, `r format_output(sum(recently_collected$gom_in_dbs$min_year > 2001)/spp_not_in_lists$gom_not_in_list * 100)`% of the species not on the checklist),  have been collected for the first time since the publication of the checklist, but more significantly for the Pacific Northwest with  `r sum(recently_collected$pnw_in_dbs$min_year > 1996)` species (`r format_output(sum(recently_collected$pnw_in_dbs$min_year > 1996)/n_distinct(total_pnw_records$worms_valid_name) * 100)`% of the total species richness,  `r format_output(sum(recently_collected$pnw_in_dbs$min_year > 1996)/spp_not_in_lists$pnw_not_in_list * 100)`% of the species not on the checklist).

>>>>

Several factors may contribute to species being included in the regional lists but not having records in biodiversity databases. The first category of biases is linked to the likelihood of having specimens for these species in museum collections. For instance, rare species may never have been collected or may have only been collected a few times, and the corresponding records have not been digitized yet. Species that are difficult to preserve because they are large, very small, or fragile, are deposited in museums less frequently. The other category of biases is associated with data quality for these records. Some of these species may be in the databases but do not have geographic coordinates associated with them, and therefore would not be included in the results of our query. While it is difficult to assess biases associated with the likelihood of preservation of a species, we investigated how missing geographic coordinates are affecting the discrepancies between the checklists and the databases.

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio, we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for those that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_usa*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_usa, big.mark=",")`) fall within US waters. Additional digitization of collections, including the capture of their geographic coordinates information, is needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems.

Two factors most likely contribute to explanations of why some species are in databases but are not included in the regional the checklists: (1) the species have been documented for the first time since the publication of the checklist, and (2) the species records were not accessible or known to the authors.

Recently recorded species contribute modestly to this discrepancy for the Gulf of Mexico, as `r sum(recently_collected$gom_in_dbs$min_year > 2001)` species (`r format_output( sum(recently_collected$gom_in_dbs$min_year > 2001)/n_distinct(total_gom_records$worms_valid_name) * 100)`% of the total species richness, `r format_output(sum(recently_collected$gom_in_dbs$min_year > 2001)/spp_not_in_lists$gom_not_in_list * 100)`% of the species not on the checklist),  have been collected for the first time since the publication of the checklist, but the results are more significant for the Pacific Northwest with  `r sum(recently_collected$pnw_in_dbs$min_year > 1996)` species (`r format_output(sum(recently_collected$pnw_in_dbs$min_year > 1996)/n_distinct(total_pnw_records$worms_valid_name) * 100)`% of the total species richness,  `r format_output(sum(recently_collected$pnw_in_dbs$min_year > 1996)/spp_not_in_lists$pnw_not_in_list * 100)`% of the species not on the checklist).
